---
title: "A4_cheatsheet"
author: "Victor Møller"
date: "30 apr 2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#packages
library(pacman)
p_load(tidyverse, readxl, brms, metafor, brmstools, ggunchained, cowplot)

#your data
pitch <- read_excel("Assignment4PitchDatav2.xlsx")
meta <- read_excel("Assignment4MetaData.xlsx")
```

Transformations to be done on meta-data.
Math taken from: 
https://onlinelibrary-wiley-com.ez.statsbiblioteket.dk:12048/doi/epdf/10.1002/sim.1525

Steps in the next chunk:  
1. clean subset of meta-data. 
2. calculate the variation of standard deviation.
3. use the variation to calculate the logarithm of standard deviation.
4. use the variation to calculate the logarithm of the mean.
5. use escalc() to obtain effect size which you'll be using. 
--> yi (new varialbe) = effect size.
--> vi (new variable) = standard deviation of effect size. 
6. These two variables which escalc() outputs (yi, vi) are what you will be using instead of "MeanES" & "VarianceES". 

```{r}

#subset with clean data (I.e., only studies with the variables we need)
meta_clean <- meta %>%
  subset(MeanES != "NA")

#calculating variation (necessary to obtain log of standard deviation).
meta_clean <- meta_clean %>%
  mutate(varSDSZ = log(1+(PITCH_F0SD_SZ_SD^2/PITCH_F0SD_SZ_M^2)),
         varSDHC = log(1+(PITCH_F0SD_HC_SD^2/PITCH_F0SD_HC_M^2)))

#calculation for log of standard deviation from variance
meta_clean <- meta_clean %>%
  mutate(logSDSZ = sqrt(varSDSZ),
         logSDHC = sqrt(varSDHC))

#calculating the logarithm of mean from (old mean) & variance
meta_clean <- meta_clean %>%
  mutate(logmeanSZ = log(PITCH_F0SD_SZ_M) - 0.5 * varSDSZ,
         logmeanHC = log(PITCH_F0SD_HC_M) - 0.5 * varSDHC)

#effect size for standard deviation of pitch
meta_ready=escalc('SMD', n1i=SAMPLE_SIZE_SZ, n2i=SAMPLE_SIZE_CT, m1i=logmeanSZ,
m2i=logmeanHC, sd1i=logSDSZ, sd2i=logSDHC,
data = meta_clean)

```

now make your meta-model:

```{r}

brm_matti <- brm(
  yi | se(vi) ~ 1 + (1 | StudyRef), #works, we know the variance. 
  prior = set_prior("uniform(0, 10)", class = "sd"), 
  data = meta_ready, 
  iter = 10000,
  cores = 4
) 

brm_matti

```

steps in prep of pitch data: 
1. take the logarithm of the outcome variable.
2. z-scale it as well (to make it compatible with the meta-analysis output).

```{r}

#transforming pitch
pitch_trans <- pitch %>%
  mutate(pitchSDLOG = log(PitchSD), #logarithm
         pitchSDLOGz = scale(pitchSDLOG), #z-scaling 
         diagnosis = as.factor(diagnosis)) #as factor (not necessary)

```

now make your model

```{r}

#the regression
mod_f <- bf("pitchSDLOGz ~ diagnosis + (1|studynr) + (1 | ID) + (1|trial)")

#running the model with regularizing priors 
brm_mod <- brm(
  mod_f, 
  prior = c(
    prior(normal(0, 0.2), class = "Intercept"),
    prior(normal(0, 0.2), class = "sd"), #Rhat = 1.01, could be more regularizing.
    prior(normal(0, 0.3), class = "b"),
    prior(normal(0, 0.3), class = "sigma")),
  data = pitch_trans, 
  cores = 4, iter = 4000, warmup = 2000,
  sample_prior = TRUE #added for later use. 
)

#pp_check now looks ok. 
pp_check(brm_mod, nsamples = 100)

```

The priors from the meta-model can now be incorporated as priors into the model
from the previous chunk because we are working on the same scale. 

--------

transforming back to the natural scale
steps:
this function assumes that you have 
(1) a dataframe (dataframe1) with predicted posterior (PredictivePosterior) for the effect of diagnosis (these are the things you put in). You can use the predict() function to obtain these. 
(2) a dataframe (dataframe2) containing the logarithm of pitch standard deviation and pitch standard deviation. This dataframe will be the one which has the original observations (I.e., PitchData). 

```{r}

#the function
transfer <- function(dataframe1, PredictivePosterior) {
  dataframe1$on_log_scale <- dataframe1[[PredictivePosterior]]*sd(dataframe2$pitch_sd_log) +
    mean(dataframe2$pitch_sd_log)
  dataframe1$on_natural_scale <- exp(dataframe1$on_log_scale) 
  dataframe1$on_z_scale <- (dataframe1$on_natural_scale - mean(dataframe2$pitch_sd))/sd(dataframe2$pitch_sd)
  return(dataframe1)
}

#use the function
transfer(dataframe1, "PredictivePosterior")
```


